Title Goes Here

### 2021-02-19 analysis/paper/paper.Rmd
O3

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  cache.extra = 42,
  cache.path = "../cache/paper/",
  fig.path = "../figures/",
  dpi = 300,
  dev = c("pdf", "png")
)

@@

pattern_crosshatch <- ggplot2::ggproto("GeomCrosshatch", ggpattern::GeomPolygonPattern)

ggplot2::update_geom_defaults(pattern_crosshatch, 
                              list(pattern = "crosshatch",
                                   pattern_colour = "black",
                                   pattern_fill = "black",
                                   pattern_density = 0.01,
                                   pattern_alpha = 0.2,
                                   pattern_spacing = 0.02))

@@

for (l in seq_along(levs)[-1]) {
  cor <- sams[lev %in% levs[seq(l - 1, l)]] %>% 
    .[term == "full"] %>%
    dcast(time ~ lev, valuev.ar = "estimate") %>% 
    setnames(., colnames(.), c("time", "old", "new")) %>% 
    .[, cor(old, new)]

@@

@@

rotations_pc1 <- lapply(angles, function(a) {
  ceof$eof[[1]]$left[PC == "PC1"] %>%
    copy() %>%
    .[, hgt := rotate(hgt, a)] %>%
    sep_ReIm() %>%
    .[, .(mean = mean(hgt)), by = part] %>%
    .[, rotation := a]
}) %>%
  rbindlist()

best_rotation_pc1 <- rotations_pc1[part == "Real"][which.min(mean)]$rotation

@@

(ref:extended-series-cap) Temporal series extended using ERA5 back extended preliminary edition (period 1950 -- 1978) and ERA5 (period 1979 -- 2019). Each series is computed by projecting monthly geopotential height zonal anomalies standardised by level south of 20ºS onto the corresponding spatial pattern.

@@

era_more %>%
  copy() %>% 
  # .[, hgt := as.numeric(scale(hgt)), by = .(part, PC)] %>%
  ggplot(aes(time, hgt)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt), by  = .(PC, part)], 
             aes(yintercept = V1), size = 0.1, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  geom_line() +
  scale_alpha_manual(values = c(1, 0.2)) +

@@

```{r compute-regression}
regr <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[season == "SON"] %>%
  sep_ReIm(format = "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(R = as.numeric(scale(R)), I = as.numeric(scale(I))),
    by = .(season, PC)] %>%
  dcast(time ~ PC, value.var = c("R", "I")) %>%
  .[melt(era5, id.vars = c("time", "lon", "lat", "lev")), on = "time", allow.cartesian = TRUE] %>%
  .[lev %in% c(850, 700, 200, 50)] %>%
  # .[variable == "hgt"] %>%
  .[, FitLm(value, I_PC1, R_PC1, I_PC2, R_PC2, se = TRUE),
    by = .(lev, lon, lat, variable)] %>%
  .[, c("term", "PC") := tstrsplit(term, "_")] %>%
  rm_intercept() %>%
  .[, term := factor(term, levels = c("R", "I"), labels = c("Real", "Imaginary"))]
```

@@

  if ("lev" %in% colnames(data)) {
    data <- data[lev %in% which_levs]
  } else {
    data[, lev := which_levs]
  }
  data[lat <= -20] %>%
    # .[term == "I" & lev == 50] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), by = .(lev, PC, term)] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    ditto(
      geom_contour_fill(aes(z = estimate), breaks = AnchorBreaks(exclude = 0), global.breaks = FALSE) +
        geom_contour_tanaka() +
        scale_fill_divergent(var_title, guide = guide_colorstrip_bottom())
    ) +

    geom_contour2(aes(z = p_val), breaks = 0.05, size = 0.1) +
    stat_subset(aes(subset = p_val <= 0.05 & is.cross(lon, lat, 1)),
                geom = "point", size = 0.2, alpha = 0.1) +
    geom_qmap(~ .x[lat <= -20]) +

@@

@@

```{r}
pc2 <- ceof$eof[[1]]$left[PC == "PC2"] %>% 
  copy() %>% 
  .[, PC := NULL] %>% 
  .[]

@@

pc2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  .[, hgt := rotate(hgt, psa_angle)] %>%
  sep_ReIm() %>% 
  .[] %>% 
  ggplot(aes(value, hgt)) +
  geom_point() +
  facet_grid(part ~ PC)
```

@@

```{r eof-filtered, fig.cap = "Spatial pattern of the leading EOF of 200 hPa geopotential height with the variability of EOF2 filtered out South of 20º. Arbitrary scale."}
eof_filtered <- era5 %>% 
  .[sep_ReIm(pc2, format = "wide")[, hgt := NULL], on = .NATURAL] %>% 
  .[lev == 200] %>% 
  .[lat <= -20] %>% 
  .[season(time) == "SON"] %>% 
  .[, hgt := resid(.lm.fit(cbind(1, R, I), hgt)), .(lon, lat, lev, season(time))] %>%
  .[, eof := Anomaly(hgt)*sqrt(cos(lat*pi/180)), by = .(lon, lat, lev, season(time))] %>% 
  EOF(eof ~ lon + lat | time, n = 1, data = .)

@@

```{r regr_psi}
regr_psi <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(R = R/sd(R), I = I/sd(I)), by = .(season, PC)] %>%
  .[psi, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(psi.z, R, I, se = TRUE),by = .(lon, lat, lev, PC, season)] %>%
  rm_intercept() %>%
  .[, psi.z := Anomaly(estimate), by = .(PC, season, lev, lat)] %>%
  .[, c("fx", "fy") := shceof::WaveFlux(.SD, p = 200), by = .(lev, PC, season, term)] %>%
  .[, term := factor(term, levels = c("R", "I"), labels = c("Real", "Imaginary"),
                     ordered = TRUE)]
```

@@

```{r regr_sst}
regr_sst <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(R = R/sd(R), I = I/sd(I)), by = .(season, PC)] %>%
  .[sst, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(t, I, R, se = TRUE), by = .(lon, lat, PC, season)] %>%
  rm_intercept() %>%
  .[, term := factor(term, levels = c("R", "I"), labels = c("Real", "Imaginary"),
                     ordered = TRUE)]
```

@@

  regr_psi %>%
    .[PC == which_pc] %>%
    .[lev == which_lev] %>%
    # .[abs(estimate) > 1, estimate := NA] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "fdr"), by = .(PC, term)] %>%
    .[, var := "Streamfunction"] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e8, fill = ..level..),
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour2(aes(z = p_val), breaks = 0.05, size = 0.1) +
    stat_subset(aes(subset = p_val <= 0.05 & is.cross(lon, lat, 1)),
                geom = "point", size = 0.2, alpha = 0.5) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10, n = 30,
                    size = 0.2, res = 2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised("Streamfunction", guide = "none") +
    ggnewscale::new_scale_fill() +

@@

```{r regr_pp}
regr_pp <- ceof[, eof[[1]]$left, by = season] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(R = R/sd(R), I = I/sd(I)), by = .(season, PC)] %>%
  .[melt(cmap, id.vars = c("lon", "lat", "time")), on = "time", allow.cartesian = TRUE] %>%
  .[season == "SON"] %>%
  .[, FitLm(value, I, R, se = TRUE), by = .(lon, lat, PC, variable, season)] %>%
  rm_intercept()
```

@@

```{r plot_pp}
plot_pp <- function(data, which_pc, which_term = c("R", "I"), which_continent = NULL) {
  filter_continent <- function(data) {
    if (!is.null(which_continent)) {
      data[ continent == which_continent]
    } else {
      data
    }

  }
  data %>%
    .[PC %in% which_pc] %>%
    .[term %in% which_term] %>%
    .[, lev := "sfc"] %>%
    .[lat >= -60] %>%
    add_continent() %>%
    filter_continent() %>%
    # .[abs(estimate) <= 2] %>%
    .[, p_val := Pvaluate(estimate, std.error, df, "none"), by = .(lev, PC, term)] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    # geom_raster(aes(fill = estimate)) +
    geom_contour_fill(aes(z = estimate, fill = ..level..), 
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(exclude = 0)) +

    # scale_fill_divergent() +
    scale_fill_divergent_discretised(NULL, 
                         # limits = c(-.8, .8),
                         # oob = scales::squish,
                         low = "#8E521C",
                         high = "#00665E",
                         guide = guide_colorsteps_bottom(15)) +
    # scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(15)) +
    geom_contour2(aes(z = p_val), breaks = 0.05, size = 0.1) +
    stat_subset(aes(subset = p_val <= 0.05 & is.cross(lon, lat)),
                geom = "point", size = 0.2) +
    geom_qmap(~.x %>% .[lat %between% c(-60, max(data$lat))] %>% add_continent() %>% filter_continent(),
              keep = 1) +

    scale_x_longitude(ticks = 15) +
    scale_y_latitude(ticks = 15) +
    ggh4x::facet_nested(term ~ PC, labeller = labeller(term = lab_cplx)) +
    axis_labs_smol +
    tag_facets("cr") +
    coord_quickmap()
}
```

@@

rotations_pc1 %>%
  ggplot(aes(rotation, mean)) +
  geom_vline(xintercept = best_rotation_pc1) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line(aes(color = part), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Mean value") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```

>>>>

```{r setup, echo = FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  cache = TRUE,
  comment = "#>",
  cache.extra = 40,
  cache.path = "../cache/paper/",
  fig.path = "../figures/",
  dpi = 300,
  dev = c("pdf", "png")
)

@@

pattern_dots <- ggplot2::ggproto("GeomDots", ggpattern::GeomPolygonPattern)
ggplot2::update_geom_defaults(pattern_dots, 
                              list(pattern = "circle",
                                   colour = NA,
                                   pattern_colour = "black",
                                   pattern_fill = "black",
                                   pattern_density = 0.1,
                                   pattern_alpha = 1,
                                   fill = NA,
                                   pattern_spacing = 0.02))

geom_contour_pval <- function(mapping, p.value = 0.01, size = 0.1) {
  list(
    stat_contour2(mapping, breaks = p.value, geom = pattern_dots),
    geom_contour2(mapping, breaks = p.value, size = size)
  )
}

@@

for (l in seq_along(levs)[-1]) {
  cor <- sams[lev %in% levs[seq(l - 1, l)]] %>% 
    .[term == "full"] %>%
    dcast(time ~ lev, value.var = "estimate") %>% 
    setnames(., colnames(.), c("time", "old", "new")) %>% 
    .[, cor(old, new)]

@@

```{r o3}
o3 <- O3() %>% 
  readRDS() %>% 
  .[lat <= -20] %>% 
  .[, .(toc = mean(total_ozone_column, na.rm = TRUE)), by = .(lon, lat, time = seasonally(time))] 
o3[, time := year(time)]
```

@@

# rotations_pc1 <- lapply(angles, function(a) {
#   ceof$eof[[1]]$left[PC == "PC1"] %>%
#     copy() %>%
#     .[, hgt := rotate(hgt, a)] %>%
#     sep_ReIm() %>%
#     .[, .(mean = mean(hgt)), by = part] %>%
#     .[, rotation := a]
# }) %>%
#   rbindlist()
# 
# 
# best_rotation_pc1 <- rotations_pc1[part == "Real"][which.min(mean)]$rotation

wave1_o3 <- o3 %>% 
  .[lat %between% c(-70, -45)] %>% 
  .[, .(toc = weighted.mean(toc, w = cos(lat*pi/180), na.rm = TRUE)), by = .(time, lon)] %>% 
  .[, FitWave(toc, 1), by = .(time)]

with_o3 <- ceof$eof[[1]]$left %>% 
  .[PC == "PC1"] %>% 
  .[, time := year(time)] %>% 
  .[wave1_o3, on = "time"]

rotations_o3 <- lapply(angles, function(a) {
  with_o3 %>% 
    copy() %>% 
    .[, hgt := rotate(hgt, a)] %>% 
    na.omit() %>% 
    sep_ReIm() %>% 
    .[, .(cor = cor(amplitude, hgt)), by = part] %>% 
    .[, rotation := a]
}) %>% 
  rbindlist()

best_rotation_pc1 <- rotations_o3[part == "Real"][which.max(cor)]$rotation

@@

(ref:extended-series-cap) Temporal series extended using ERA5 back extended preliminary edition (period 1950 -- 1978) and ERA5 (period 1979 -- 2019). Each series is computed by projecting monthly geopotential height zonal anomalies standardised by level south of 20ºS onto the corresponding spatial pattern and scaled to zero mean and unit standard deviation.

@@

# era_less <-   ceof %>%
#   .[, eof[[1]]$left, by = season] %>%
#   sep_ReIm(hgt) %>%
#   .[season == "SON"] %>% 
#   .[, hgt := as.numeric(scale(hgt)), by = .(part, PC)]

era_more %>%
  copy() %>% 
  .[, ref := year(time) > 1979] %>% 
  .[, hgt := scale(hgt), by = .(part, PC)] %>%
  ggplot(aes(time, hgt)) +
  # Zero line + rangeframe
  geom_hline(data = ~.x[, mean(hgt), by  = .(PC, part)], 
             aes(yintercept = V1), size = 0.1, color = "gray30") +
  ggthemes::geom_rangeframe(sides = "lr", size = 0.1, color = "gray30") +
  geom_line() +
  # geom_line(data = era_less) +
  scale_alpha_manual(values = c(1, 0.2)) +

@@

```{r compute-regression}
regr <- ceof[, eof[[1]]$left, by = .(season)] %>%
  .[season == "SON"] %>%
  sep_ReIm(format = "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = as.numeric(scale(Real)), Imaginary = as.numeric(scale(Imaginary))),
    by = .(season, PC)] %>%
  dcast(time ~ PC, value.var = c("Real", "Imaginary")) %>%
  .[melt(era5, id.vars = c("time", "lon", "lat", "lev")), on = "time", allow.cartesian = TRUE] %>%
  .[lev %in% c(850, 700, 200, 50)] %>%
  # .[variable == "hgt"] %>%
  .[, FitLm(value, Imaginary_PC1, Real_PC1, Imaginary_PC2, Real_PC2, se = TRUE),
    by = .(lev, lon, lat, variable)] %>%
  .[, c("term", "PC") := tstrsplit(term, "_")] %>%
  rm_intercept() %>%
  .[, term := factor_ReIm(term)]
```

@@

  if ("lev" %in% colnames(data)) {
    data <- data[lev %in% which_levs]
  } else {
    data[, lev := which_levs]
  }
  data[lat <= -20] %>%
    # .[term == "I" & lev == 50] %>%
    .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(lev, PC, term)] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    ditto(
      geom_contour_fill(aes(z = estimate), breaks = AnchorBreaks(exclude = 0), global.breaks = FALSE) +
        geom_contour_tanaka() +
        scale_fill_divergent(var_title, guide = guide_colorstrip_bottom())
    ) +
    geom_contour_pval(aes(z = p.value)) +

    geom_qmap(~ .x[lat <= -20]) +

@@

```{r t}
t <- ERA5() %>% 
  ReadNetCDF(vars = "t", 
             subset = list(latitude = c(-90, -20))) %>% 
  normalise_coords()
```

```{r t-vertical, fig.cap = "Regression of EOF1 with mean air temperature between 80ºS and 25ºS."}
# t_PC1_vertical <- t %>% 
#   .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
#   sep_ReIm(format = "wider") %>% 
#   .[, FitLm(t, Real, Imaginary, se = TRUE), by = .(lev, lon, lat)] %>% 
#   rm_intercept()
# t_PC1_vertical %>% 
#   .[lev %in% c(1, 5, 10, 50)] %>% 
#   .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(term, lev)] %>% 
#     ggplot(aes(lon, lat)) +
#   geom_contour_fill(aes(z = estimate)) +
#   stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)),
#               size = .5, alpha = 0.5) +
#   geom_hline(yintercept = c(-80, -35)) +
#   geom_qmap(~.x[lat <= -20]) +
#   scale_fill_divergent() + 
#   coord_polar() +
#   facet_grid(lev ~ term)

t %>% 
  .[lat %between% c(-80, -35)] %>% 
  .[, .(t = weighted.mean(t, cos(lat*pi/180))), by = .(lev, time, lon)] %>%  
  .[cut(ceof$eof[[1]], 1)$left, on = "time"] %>% 
  sep_ReIm(format = "wider") %>% 
  .[, FitLm(t, Real, Imaginary, se = TRUE), by = .(lev, lon)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)] %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(term)] %>%
  ggplot(aes(lon, lev)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_tanaka(aes(z = estimate), 
                      breaks = AnchorBreaks(0, exclude = 0)) +

  geom_contour_pval(aes(z = p.value)) +
  scale_fill_divergent_discretised("Air temperature", guide = guide_colorsteps_bottom()) + 
  scale_y_level() +
  scale_x_longitude() +
  facet_grid(~term) +
  axis_labs_smol
```

```{r o3_regr} 
temp <- ceof$eof[[1]]$left %>% 
  .[PC == "PC1"] %>% 
    copy() %>% 
  sep_ReIm(format = "wider") %>% 
  .[, `:=`(Real = as.numeric(scale(Real)), 
           Imaginary = as.numeric(scale(Imaginary))),
    by = .(PC)] %>%
  .[, time := year(time)]

o3_regr <- temp %>% 
  o3[., on = "time", allow.cartesian = TRUE] %>% 
  .[, FitLm(toc, Real, Imaginary, se = TRUE), by = .(lon, lat, PC)] %>% 
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)]
```

```{r o3-regr, fig.cap = "Regression of standarised EOF1 and Total Ozone column."}
o3_regr %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(PC, term)] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate, fill = ..level..),
                    breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(0, exclude = 0)) +
  geom_contour_pval(aes(z = p.value)) +
  geom_qmap(~.x[lat <= -20]) +
  scale_fill_divergent_discretised("Total Ozone Column", 
                                   guide = guide_colorsteps_bottom()) + 
  scale_x_longitude() +
  scale_y_latitude() +
  coord_polar() +
  facet_grid(PC~term) +
  axis_labs_smol
```

The ozone hole, as measured by Total Ozone Column (TOC) is climatologically not centered over the South Pole, but shifted towards the Weddell Sea, which translates to a large amplitude zonal wave 1 with minimum over the Weddell Sea and maximum over the Davis Sea [@wirth1993]. The regression

```{r wave1-o3, fig.cap = "EOF1 and wave1 o3 amplitude."}
wave1_o3 <- o3 %>% 
  .[lat %between% c(-70, -45)] %>% 
  .[, .(toc = mean(toc, na.rm = TRUE)), by = .(time, lon)] %>% 
  .[, FitWave(toc, 1), by = .(time)]

temp <- ceof$eof[[1]]$left %>% 
  .[PC == "PC1"] %>% 
  .[, time := year(time)] %>% 
  .[wave1_o3, on = "time"]

temp %>% 
  copy() %>% 
  na.omit() %>% 
  sep_ReIm() %>% 
  .[, hgt := scale(hgt), by  = .(part)] %>% 
  ggplot(aes(amplitude, hgt)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(PC ~ part)
```

@@

```{r psa-eof2, fig.cap = "EFO2 and PSA"}
pc2 <- ceof$eof[[1]]$left[PC == "PC2"] %>% 
  copy() %>% 
  .[, PC := NULL] %>% 
  .[]

@@

pc2[psa_time, on = "time"] %>% 
  .[PC != "SAM"] %>% 
  copy() %>% 
  .[, hgt := rotate(hgt, psa_angle)] %>%
  sep_ReIm() %>% 
  ggplot(aes(value, hgt)) +
  geom_point() +
  scale_y_continuous("EOF2") +
  scale_x_continuous("PSA") +
  facet_grid(part ~ PC)
```

@@

```{r eof-filtered, fig.cap = "Spatial pattern of the leading EOF of 200 hPa geopotential height with the variability of EOF2 filtered out South of 20º. Arbitrary scale."}
eof_filtered <- era5 %>% 
  .[sep_ReIm(pc2, format = "wide")[, hgt := NULL], on = .NATURAL] %>% 
  .[lev == 200] %>% 
  .[lat <= -20] %>% 
  .[season(time) == "SON"] %>% 
  .[, hgt := resid(.lm.fit(cbind(1, Real, Imaginary), hgt)), .(lon, lat, lev, season(time))] %>%
  .[, eof := Anomaly(hgt)*sqrt(cos(lat*pi/180)), by = .(lon, lat, lev, season(time))] %>% 
  EOF(eof ~ lon + lat | time, n = 1, data = .)

@@

```{r regr_psi}
regr_psi <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), by = .(season, PC)] %>%
  .[psi, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(psi.z, Real, Imaginary, se = TRUE),by = .(lon, lat, lev, PC, season)] %>%
  rm_intercept() %>%
  .[, psi.z := Anomaly(estimate), by = .(PC, season, lev, lat)] %>%
  .[, c("fx", "fy") := shceof::WaveFlux(.SD, p = 200), by = .(lev, PC, season, term)] %>%
  .[, term := factor_ReIm(term)]
```

@@

```{r regr_sst}
regr_sst <- ceof[, eof[[1]]$left, by = .(season)] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), by = .(season, PC)] %>%
  .[sst, on = "time", allow.cartesian = TRUE] %>%
  .[season(time) == "SON"] %>%
  .[, FitLm(t, Imaginary, Real, se = TRUE), by = .(lon, lat, PC, season)] %>%
  rm_intercept() %>%
  .[, term := factor_ReIm(term)]
```

@@

  regr_psi %>%
    .[PC == which_pc] %>%
    .[lev == which_lev] %>%
    # .[abs(estimate) > 1, estimate := NA] %>%
    .[lat %between% c(-90, 10)] %>%
    .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(PC, term)] %>%
    .[, var := "Streamfunction"] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    geom_contour_fill(aes(z = estimate*1e8, fill = ..level..),
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_pval(aes(z = p.value)) +
    geom_streamline(aes(dx = fx, dy = fy), min.L = 3, L = 10, n = 30,
                    size = 0.2, res = 2) +
    geom_qmap(keep = .015) +
    scale_fill_divergent_discretised("Streamfunction", guide = "none") +
    ggnewscale::new_scale_fill() +

@@

```{r regr_pp}
regr_pp <- ceof[, eof[[1]]$left, by = season] %>%
  sep_ReIm(hgt, "wide") %>%
  .[, hgt := NULL] %>%
  .[, `:=`(Real = Real/sd(Real), Imaginary = Imaginary/sd(Imaginary)), by = .(season, PC)] %>%
  .[melt(cmap, id.vars = c("lon", "lat", "time")), on = "time", allow.cartesian = TRUE] %>%
  .[season == "SON"] %>%
  .[, FitLm(value, Imaginary, Real, se = TRUE), by = .(lon, lat, PC, variable, season)] %>%
  rm_intercept() %>% 
  .[, term := factor_ReIm(term)]
```

@@

```{r plot_pp}
plot_pp <- function(data, which_pc, which_term = c("Real", "Imaginary"), which_continent = NULL) {
  filter_continent <- function(data) {
    if (!is.null(which_continent)) {
      data[continent == which_continent]
    } else {
      data
    }

  }
  data %>%
    .[PC %in% which_pc] %>%
    .[term %in% which_term] %>%
    .[, lev := "sfc"] %>%
    .[lat >= -60] %>%
    add_continent() %>%
    filter_continent() %>%
    # .[abs(estimate) <= 2] %>%
    .[, p.value := Pvaluate(estimate, std.error, df, "none"), by = .(lev, PC, term)] %>%
    periodic(lon = c(0, 360)) %>%
    ggplot(aes(lon, lat)) +
    # geom_raster(aes(fill = estimate)) +
    geom_contour_fill(aes(z = estimate, fill = ..level..), 
                      breaks = AnchorBreaks(exclude = 0)) +
    geom_contour_tanaka(aes(z = estimate), breaks = AnchorBreaks(exclude = 0)) +

    # scale_fill_divergent() +
    scale_fill_divergent_discretised(NULL, 
                         # limits = c(-.8, .8),
                         # oob = scales::squish,
                         low = "#8E521C",
                         high = "#00665E",
                         guide = guide_colorsteps_bottom(15)) +
    # scale_fill_divergent_discretised(NULL, guide = guide_colorsteps_bottom(15)) +
    geom_contour_pval(aes(z = p.value)) +
    geom_qmap(~.x %>% .[lat %between% c(-60, max(data$lat))] %>% add_continent() %>% filter_continent(),
              keep = 1) +

    scale_x_longitude(ticks = 15) +
    scale_y_latitude(ticks = 15) +
    ggh4x::facet_nested(term ~ PC) +
    axis_labs_smol +
    tag_facets("cr") +
    coord_quickmap()
}
```

@@

rotations_o3 %>%
  ggplot(aes(rotation, cor)) +
  geom_vline(xintercept = best_rotation_pc1) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_line(aes(color = part), size = 1) +
  scale_x_continuous("Rotation", breaks = seq(-pi, pi, by = 30*pi/180),
                     labels = function(x) paste0(round(x*180/pi, 5), "º")) +
  scale_y_continuous("Correlation with O3 zonal wave 1") +
  scale_color_brewer(NULL, palette = "Dark2", labels = lab_cplx)
```
